Numerical investigation of viscous flow instabilities through heterogeneous porous media. Advantages and limitations of unstructured and adaptive meshes in 2D and 3D domains under different mobility scenarios.

### 2017-12-07 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

%%%%%%%%%%%%%
\subsection{3-D channel flows} \label{section:results_3D}

In this section, 3D simulations are introduced to investigate preferential flow pathways and finger growth through semi-pervious and impervious layers. In large scale geological formations preferential flow may allow fast contaminant transport (therefore creating significant consequences for groundwater quality) and low recovery rate (in hydrocarbon production fields).

As outlined in Section \ref{section:ViscousInstabilities}, viscous instabilities are characterised by spontaneous formation and growth of fluid elongations, called viscous fingers. The geometrical evolution of these elongations provides good indication of instabilities at the fluid front interface. Figure \ref{fig:3DChannel_PermMesh} shows initial permeability distribution for the simulations described in this section. The surface on the left-hand side of the domain is the area where fluid $1$ is injected with a uniform velocity of 2.5 cm.s$^{-1}$. No-flux boundary conditions are applied to all borders, except the right-hand side (inlet) and the left-hand side (outflow). Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0)=0.80$, which was displaced by a pure (wetting) fluid. The domain (with 368513 \PNDG[1]{1} tetrahedral element-pairs) was designed to capture preferential flow pathways, \ie regions with larger permeability (Fig.~\ref{fig:3DChannel_PermMesh}) as demonstrated in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.

>>>>

%%%%%%%%%%%%%
\subsection{3-D channel flows} \label{section:results_3D} 
In large scale geological formations, preferential flow may result in fast polutant transport (advection-diffusion-dispersion) after industrial spillage (therefore leading to contamination of groundwater) and low hydrocarbon recovery rates (in oil/gas exploration fields).  In this section, 3D simulations are introduced to investigate preferential flow pathways through semi-pervious and impervious geological layers.

%As outlined in Section \ref{section:ViscousInstabilities}, viscous instabilities are characterised by spontaneous formation and growth of fluid elongations, called viscous fingers. The geometrical evolution of these elongations provides good indication of instabilities at the fluid front interface.
Figure \ref{fig:3DChannel_PermMesh} shows initial permeability distribution for the simulations described in this section. The surface on the left-hand side of the domain is the area where fluid $1$ is injected with a uniform velocity of 2.5 cm.s$^{-1}$. No-flux boundary conditions are applied to all borders, except the right-hand side (inlet) and the left-hand side (outflow). Initially, the domain was partially saturated with non-wetting fluid, \ie $S_{nw}(t=0)=0.80$, which was displaced by a pure (wetting) fluid. The domain (with 368513 \PNDG[1]{1} tetrahedral element-pairs) was designed to capture preferential flow pathways, \ie regions with larger permeability (Fig.~\ref{fig:3DChannel_PermMesh}) as demonstrated in Fig.~\ref{fig:3DChannel_planes}-\ref{fig:3DChannel_Isosurf}.
